---
title: "Personal Activity analyzer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#Loading and preprocessing the data

```{r }
setwd("C:\\Coursera\\DATA_SCIENCE\\Reproducible Research")
getwd()

zipF1<- "C:\\Coursera\\DATA_SCIENCE\\Reproducible Research\\activity.zip"
ractivity <- read.csv("activity.csv", header = TRUE)
```
#What is mean total number of steps taken per day?


```{r}
sums <- tapply(ractivity$steps, ractivity$date, sum)


hist(sums, breaks = 6, col = "green", xlab = "Steps", 
     main = "Total steps")
#Mean value
mean(sums,na.rm = TRUE)
#Median value
median(sums, na.rm = TRUE)

```
#What is the average daily activity pattern?
Making a time series plot 

```{r}

means <- tapply(ractivity$steps, ractivity$interval, mean, na.rm = TRUE, simplify = T)

averages <- as.numeric(names(means))

plot(averages,means, type  = "l", col = "blue", xlab = "5-minute interval", ylab = "average steps across all the days")
```

#Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}

maxval <- sort(means, decreasing = TRUE)
maxval[1]
#max number of step 206.17 on the 835th interval
```


#Imputing missing values
Creating new data set without NA values


What is the average daily activity pattern?
Make a time series plot 

```{r}

newdata <- ractivity

for(i in 1:ncol(newdata)){
  newdata[is.na(newdata[,i]),i] <- mean(newdata[,i], na.rm = TRUE)
}
```
#Number of missing values between old data set and new
```{r}
sum(is.na(ractivity))
sum(is.na(newdata))
head(ractivity)
head(newdata)


newstat <- tapply(as.numeric(newdata$steps), newdata$date,sum)
hist(newstat, main = "Steps per day without missing values")
mean(newstat, na.rm = TRUE)
median(newstat, na.rm = TRUE)
```
#Are there differences in activity patterns between weekdays and weekends?

Adding a column with days
```{r}
newdata$Week_Day<- format(as.Date(newdata$date), format = "%A")
```
Weekday activity
```{r}
weekday <- subset(newdata,newdata$Week_Day != c("Sunday","Saturday"))
head(weekday)
tail(weekday)
```

Weekend activity
```{r}
weekend <- subset(newdata,newdata$Week_Day == c("Sunday","Saturday"))
head(weekend)
tail(weekend)
```
Weekday plot
```{r}
meanweekday <- tapply(weekday$steps, weekday$interval, mean, na.rm = TRUE) 
meanweekday1 <- as.numeric(names(meanweekday))
plot (meanweekday1, meanweekday, type = "l", col = "red", main = "Weekday activity" )
```
Weekend Plot
```{r}
meanweekend <- tapply(weekend$steps, weekend$interval, mean, na.rm = TRUE) 
meanweekend1 <- as.numeric(names(meanweekend))
plot (meanweekend1, meanweekend, type = "l", col = "blue",main = "Weekend activity"  )

```

